### 2011-03-18 extra/sliver_pldi11/response.txt

We thank all reviewers for their useful comments; we will certainly
incorporate them into future versions of this paper.

REVIEWERS 1 and 4 ask about running time.  The relationship between the number
of input tuples (which we report as the main evaluation metric) and running
time is indeed obscured by using BDDs, but there is a definite correlation.  In
particular, pruning offers modest improvements in performance for small k
values, but the main win is that pruning allows us to even reach higher k
values at all (without pruning, the analysis just runs out of memory before
getting there).

REVIEWER 1: In Table 4, column k corresponds to using a k-limited abstraction.
We will improve the presentation of the results.

REVIEWER 4: The budget of the PR algorithm is determined by the number of
iterations.  In our experiments, we increased k until our analysis ran out
of memory (the horizontal axis of Figure 10 shows the k values which were
reached).

REVIEWER 4: to relate this paper with past work: Indeed, at a high-level, both
our paper and Sridharan-Bodik PLDI'06 refine an abstraction based on a client,
but there are some important differences: (i) their work is specific to pointer
analysis, whereas ours applies generally to any Datalog program; (ii) we also
provide rigorous theoretical guarantees on the precision of our pruning
approach.  As for Liang-Tripp-Naik POPL'11: (i) they focused on finding the
coarsest abstraction to prove a query by searching through the space of
abstractions---this was a scientific study, not a usable algorithm, whereas the
PR algorithm presented in this paper is very practical; (ii) we also are
effectively working with a much richer family of abstractions, where the
context-sensitivity can depend on the specific call chain rather than just the
allocation site.

REVIEWER 2:

To provide more motivation for the particular abstractions we used: First, we
worked with k-limited and class-based abstractions because these are standard
abstractions in pointer analysis which have traditionally faced scalability
issues.  On \alpha_i and \beta_i: usually \beta_i is \alpha_i composed with
another abstraction \tau.  One intuition is that \tau and \alpha_i should be
"complementary"; the effort spent pre-pruning using \tau should neither be
trivial nor redundant with \alpha_i.

Regarding how the correctness of Theorem 2 is related to soundness and
completeness: what we mean is any query we can prove directly using a fine
abstraction we can also prove with pruning (completeness) and vice-versa
(soundness).  When the fine abstraction is no abstraction (concrete), this is
the usual notion of soundness/completeness.

Figure 3 only shows computation of P(X) in the first iteration of the pruning
algorithm.  In this iteration, the red tuples pruned.  In the next iteration,
using a finer abstraction, common(G1,G2,3) will not be reachable, as desired.

REVIEWER 3:

Regarding the comment about system approximation and overapproximation: we do
not view pruning as system approximation but rather as program slicing: one
needs to be careful about what is pruned/sliced out to ensure soundness; for
example, if you prune away all the tuples, you can "prove" anything.

### 2011-03-18 extra/sliver_pldi11/abstractions.tex
fixed several typos

Recall from \refsec{preliminaries} that we separate the Datalog program that
computes the query from the abstraction.  In \refsec{klimited}, we described
the Datalog programs that computed the queries for our three clients based on
$\infty$-object-sensitivity.  We now describe the abstractions that use to
further abstract this analysis.

@@

The {\em $k$-limited abstraction}
essentially partitions chains based on their length $k$ prefix.
We first setup some notation.
%Let $\dom$ be a set of values (e.g., $\dom = \{ 0, 1 \}$).
For a chain $c \in \dom^*$, let $|c|$ denote the length of the chain.
Let $c[i]$ be the $i$-th element of $c$ (starting with index 1) and let
let $c[i..j]$ be the subchain $[c[i] \cdots c[j]]$
(boundary cases: $c[i..j] = [\,]$ if $i > j$ and $c[i..j] = c[i..|c|]$ if $j > |c|$).
For two chains $c_1, c_2$, let $c_1+c_2$ denote concatenation.
For a chain $c$, let $c*$ denote the set of all chains
with prefix $c$; formally:
\begin{align}
c* \eqdef \{ c' \times \H^* : c'[1..|c|] = c \}.
\end{align}
%It is important that $c*$ is a set.

@@

Traditionally, $k$-limited analyses
simply use a list of sites ``$c$'' to represent an abstract value,
not making a distinction between $\{c\}$ and $c*$;
however when we do pruning, this distinction is crucial.
Suppose that $k=3$.
Na\"ively, ``$\vec{01}$'' would just represent $\{\vec{01}\}$.
The problem comes when we prune.
If we prune ``$\vec{010}$'', then ``$\vec{01}$'' really represents $\{\vec{01}\}\cup \vec{010}*$
because any chain starting with $\vec{010}$ must now fall back on ``$\vec{01}$''.
This is unacceptable because pruning tuples should not change the abstraction.
Actually, the situation is worse, because (i) in fact no pruning
(in the sense that some chains are ignored)
would happen and (ii) the resulting analysis would be much more imprecise because
``$[\,]$'' is always be present for representing the context of the main method,
but this is a prefix of any chain.

@@

Then we define the {\em barely-repeating $k$-limited abstraction} $\klimdabs_k$ as follows:
\begin{align}
\label{eqn:barelyAbstraction}
\klimdabs_k(c) & \eqdef \klimabs_{\min\{k,\delta(c)\}}(c),
\end{align}
\reffig{barelyExample} shows an example of $\klimdabs_k$.
It is not automatic that $\klimabs_k$ is a valid abstraction,
but we can check that it is:
\begin{proposition}
\label{prop:barelyAbstraction}
The function $\klimdabs_k$ defined in \refeqn{barelyAbstraction} is a valid abstraction
(\refdef{abstraction}).
\end{proposition}
\begin{proof}
We consider two cases: (i) for $\{c\} \in \range(\klimdabs_k)$,
we have $\klimdabs_k(c) = \{c\}$; and (ii) for any $c* \in \range(\klimdabs_k)$,
either $|c| = k$ or $c$ is barely-repeating; in either case,
we can check that any extension $c' \in c*$ will have $\klimabs_k(c') = c*$.
\end{proof}
Remark: one might wonder why we defined the abstraction using the barely-repeating criterion
as opposed to the simpler non-repeating criterion.
It turns out that using the latter in \refeqn{barelyLen} would not result in a valid abstraction.
If $\klimdabs_k$ were defined using the non-repeating criterion,
then $\klimdabs_3(\vec{00}) = \vec{0}*$.
But for $\vec{01} \in \vec{0}*$, we have $\klimdabs_3(\vec{01}) = \{\vec{01}\} \neq \vec{0}*$.

@@

We now introduce an abstraction that we will use in the pre-pruning step of the \PR\ algorithm.
We start by defining an equivalence relation over allocation sites $\dom$,
represented by a function $\hclass : \dom \mapsto \sP(\dom)$
mapping each allocation site $h \in \dom$ to its its equivalence class.
Given such such an $\hclass$, we could then extend it sequences by elementwise application:
\begin{align}
\hclass(c) = \hclass(c[1]) \times \cdots \times \hclass(c[|c|]), \quad c \in \dom^*.
\end{align}

>>>>

Recall from \refsec{preliminaries} that we separate the Datalog program that
computes the query from the abstraction.  In \refsec{klimited}, we described
the Datalog programs that computed the queries for our three clients based on
$\infty$-object-sensitivity.  We now describe the abstractions that we use to
further abstract this analysis.

@@

The {\em $k$-limited abstraction}
essentially partitions chains based on their length $k$ prefix.
We first setup some notation.
%Let $\dom$ be a set of values (e.g., $\dom = \{ 0, 1 \}$).
For a chain $c \in \dom^*$, let $|c|$ denote the length of the chain.
Let $c[i]$ be the $i$-th element of $c$ (starting with index 1) and
let $c[i..j]$ be the subchain $[c[i] \cdots c[j]]$
(boundary cases: $c[i..j] = [\,]$ if $i > j$ and $c[i..j] = c[i..|c|]$ if $j > |c|$).
For two chains $c_1, c_2$, let $c_1+c_2$ denote concatenation.
For a chain $c$, let $c*$ denote the set of all chains
with prefix $c$; formally:
\begin{align}
c* \eqdef \{ c' \in \H^* : c'[1..|c|] = c \}.
\end{align}
%It is important that $c*$ is a set.

@@

Traditionally, $k$-limited analyses
simply use a list of sites ``$c$'' to represent an abstract value,
not making a distinction between $\{c\}$ and $c*$;
however when we do pruning, this distinction is crucial.
Suppose that $k=3$.
Na\"ively, ``$\vec{01}$'' would just represent $\{\vec{01}\}$.
The problem comes when we prune.
If we prune ``$\vec{010}$'', then ``$\vec{01}$'' really represents $\{\vec{01}\}\cup \vec{010}*$
because any chain starting with $\vec{010}$ must now fall back on ``$\vec{01}$''.
This is unacceptable because pruning tuples should not change the abstraction.
Actually, the situation is worse, because (i) in fact no pruning
(in the sense that some chains are ignored)
would happen and (ii) the resulting analysis would be much more imprecise because
``$[\,]$'' is always present for representing the context of the main method,
but this is a prefix of any chain.

@@

Then we define the {\em barely-repeating $k$-limited abstraction} $\klimdabs_k$ as follows:
\begin{align}
\label{eqn:barelyAbstraction}
\klimdabs_k(c) & \eqdef \klimabs_{\min\{k,\delta(c)\}}(c),
\end{align}
\reffig{barelyExample} shows an example of $\klimdabs_k$.
We show that $\klimdabs_k$ is a valid abstraction:
\begin{proposition}
\label{prop:barelyAbstraction}
The function $\klimdabs_k$ defined in \refeqn{barelyAbstraction} is a valid abstraction
(\refdef{abstraction}).
\end{proposition}
\begin{proof}
We consider two cases: (i) for $\{c\} \in \range(\klimdabs_k)$,
we have $\klimdabs_k(c) = \{c\}$; and (ii) for any $c* \in \range(\klimdabs_k)$,
either $|c| = k$ or $c$ is barely-repeating; in either case,
it is easy to see that any extension $c' \in c*$ will have $\klimdabs_k(c') = c*$.
\end{proof}
Remark: one might wonder why we defined the abstraction using the barely-repeating criterion
as opposed to the simpler non-repeating criterion.
It turns out that using the latter in \refeqn{barelyLen} would not result in a valid abstraction.
If $\klimdabs_k$ were defined using the non-repeating criterion,
then $\klimdabs_3(\vec{00}) = \vec{0}*$.
But for $\vec{01} \in \vec{0}*$, we have $\klimdabs_3(\vec{01}) = \{\vec{01}\} \neq \vec{0}*$.

@@

We now introduce an abstraction that we will use in the pre-pruning step of the \PR\ algorithm.
We start by defining an equivalence relation over allocation sites $\dom$,
represented by a function $\hclass : \dom \mapsto \sP(\dom)$
mapping each allocation site $h \in \dom$ to its its equivalence class.
Given such an $\hclass$, we could then extend it sequences by elementwise application:
\begin{align}
\hclass(c) = \hclass(c[1]) \times \cdots \times \hclass(c[|c|]), \quad c \in \dom^*.
\end{align}

### 2011-03-18 extra/sliver_pldi11/experiments.tex
fixed several typos

\begin{table}
\[
\begin{array}{ll}
\text{\bf Abstractions} \\
\Klimabs = (\klimabs_k)_{k=1}^\infty   & \defn{$k$-limited abstractions \refeqn{klimitedAbstraction}} \\
\Klimdabs = (\klimdabs_k)_{k=1}^\infty & \defn{barely-repeating $k$-limited abstractions \refeqn{barelyAbstraction}} \\
\hclassIs                              & \defn{abstraction using type of allocation site} \\
\hclassHas                             & \defn{abstraction using type of containing class} \\
\hclassIsHas                           & \defn{abstraction using both types} \\
\\
\text{\bf Algorithms} \\
\Full(\alpha)         & \defn{standard analysis using abstraction $\alpha$} \\
\Site(\balpha)        & \defn{site-based refinement \cite{liang11minimal} on abstractions $\balpha$} \\
\PR(\balpha)          & \defn{\PR\ algorithm using $\balpha$, no pre-pruning} \\
\PR(\balpha,\hclass)  & \defn{\PR\ algorithm using $\balpha$, using $\balpha\circ\hclass$ to pre-prune} \\
\end{array}
\]
\caption{\label{tab:algorithms} Shows the abstractions and algorithms that we evaluated empirically.
For example,
$\PR(\Klimdabs,\hclassIsHas)$
means running the \PR-algorithm on the barely-repeating $k$-limited abstraction ($\alpha_k = \klimdabs_k$),
using a composed abstraction based on the type of an allocation site ($\hclassIs$) 
and the type of the declaring class ($\hclassHas$) to do pre-pruning
(specifically, $\beta_k = \klimdabs_k \circ \hclassIsHas$.
}
\end{table}

>>>>

\begin{table}
\[
\begin{array}{ll}
\text{\bf Abstractions} \\
\Klimabs = (\klimabs_k)_{k=1}^\infty   & \defn{$k$-limited abstractions \refeqn{klimitedAbstraction}} \\
\Klimdabs = (\klimdabs_k)_{k=1}^\infty & \defn{barely-repeating $k$-limited abstractions \refeqn{barelyAbstraction}} \\
\hclassIs                              & \defn{abstraction using type of allocation site} \\
\hclassHas                             & \defn{abstraction using type of containing class} \\
\hclassIsHas                           & \defn{abstraction using both types} \\
\\
\text{\bf Algorithms} \\
\Full(\alpha)         & \defn{standard analysis using abstraction $\alpha$} \\
\Site(\balpha)        & \defn{site-based refinement \cite{liang11minimal} on abstractions $\balpha$} \\
\PR(\balpha)          & \defn{\PR\ algorithm using $\balpha$, no pre-pruning} \\
\PR(\balpha,\hclass)  & \defn{\PR\ algorithm using $\balpha$, using $\balpha\circ\hclass$ to pre-prune} \\
\end{array}
\]
\caption{\label{tab:algorithms} Shows the abstractions and algorithms that we evaluated empirically.
For example,
$\PR(\Klimdabs,\hclassIsHas)$
means running the \PR-algorithm on the barely-repeating $k$-limited abstraction ($\alpha_k = \klimdabs_k$),
using a composed abstraction based on the type of an allocation site ($\hclassIs$) 
and the type of the declaring class ($\hclassHas$) to do pre-pruning
(specifically, $\beta_k = \klimdabs_k \circ \hclassIsHas$).
}
\end{table}

### 2011-03-18 extra/sliver_pldi11/introduction.tex
fixed several typos

% Problem: want precision, abstractions don't scale up, even with
% client-driven, etc.
Making a static analysis more precise requires increasing the complexity of the
supporting abstraction---in pointer analysis, by increasing the amount of
context/object sensitivity \cite{kcfa, kobj, MilanovaRountevRyder2005,
WhaleyLam2004, LhotakHendren2006, LhotakHendren2008}; or in model checking, by
adding more abstraction predicates \cite{graf97predicate,slam}.
However, the complexity of these analyses often grows exponentially as the
abstraction increases.
Much work has been done on curbing this exponential growth
(e.g., client-driven \cite{GuyerLin2003} and demand-driven
\cite{HeintzeTardieu2001} approaches in pointer analysis,
lazy abstraction \cite{henzinger02lazy,mcmillan06lazy} and other iterative refinement approaches in model checking).
The general theme of all these approaches is that the
complexity of the abstraction is increased only in parts deemed useful using
some feedback from a client query.

@@

% Slivers
We apply our pruning technique to the $k$-object-sensitivity abstraction
\cite{kobj}, where objects on the heap are abstracted into a chain of
allocation sites.  Pruning in this setting corresponds to maintaining a subset
of these chains.  Although the main contribution of the paper is the general
pruning technique, we also make the following contributions specific to
$k$-limited pointer analysis which arise because we are pruning: First, we show
that we need a more careful treatment of the $k$-object-sensitivity abstraction
than before due to pruning.  Second, we introduce an abstraction which limits
chains to length $k$ and also truncates to avoid having too many repeating
sites; this is an effective way to increase $k$ further (possible now due to
pruning) without getting bogged down by recursion.  Third, we show that an
abstraction that replaces allocation sites by types (a generalization of
\cite{smaragdakis11context}) is effective for pruning.

% Experiments
We ran our experiments on \numBenchmarks\ Java benchmarks using three clients that
depend heavily on having an accurate pointer analysis: downcast safety
checking, monomorphic call site inference, and race detection.
We show that with pruning, our \PR\ algorithm enables us to perform $k$-object-sensitivity
analysis with a substantially much finer abstraction (larger $k$)
compared to a full $k$-object-sensitive analysis or even using the selected
refinement strategy of \cite{liang11minimal}.
In some cases, the non-pruning approaches hit a wall around $k=3$ but the \PR\
algorithm is able to go well beyond $k=10$.

>>>>

% Problem: want precision, abstractions don't scale up, even with
% client-driven, etc.
Making a static analysis more precise requires increasing the complexity of the
supporting abstraction---in pointer analysis, by increasing the amount of
context/object sensitivity \cite{kcfa, kobj, MilanovaRountevRyder2005,
WhaleyLam2004, LhotakHendren2006, LhotakHendren2008}; or in model checking, by
adding more abstraction predicates \cite{graf97predicate,slam}.
However, the complexity of these analyses often grows exponentially as the
abstraction increases.
Much work has been done on curbing this exponential growth
(e.g., client-driven \cite{GuyerLin2003} and demand-driven
\cite{HeintzeTardieu2001} approaches in pointer analysis,
lazy abstraction \cite{henzinger02lazy,mcmillan06lazy} and other iterative refinement approaches in model checking).
The general theme of all these approaches is that the
complexity of the abstraction is increased only in parts deemed relevant using
some feedback from a client query.

@@

% Slivers
We apply our pruning technique to the $k$-object-sensitivity abstraction
\cite{kobj}, where objects in the heap are abstracted using chains of
allocation sites.  Pruning in this setting corresponds to maintaining a subset
of these chains.  Although the main contribution of the paper is the general
pruning technique, we also make the following contributions specific to
$k$-limited pointer analysis which arise because we are pruning: First, we show
that we need a more careful treatment of the $k$-object-sensitivity abstraction
than before due to pruning.  Second, we introduce an abstraction which limits
chains to length $k$ and also truncates to avoid having too many repeating
sites; this is an effective way to increase $k$ further (possible now due to
pruning) without getting bogged down by recursion.  Third, we show that an
abstraction that replaces allocation sites by types (a generalization of
\cite{smaragdakis11context}) is effective for pruning.

% Experiments
We ran our experiments on \numBenchmarks\ Java benchmarks using three clients that
depend heavily on having a precise pointer analysis: downcast safety
checking, monomorphic call site inference, and race detection.
We show that with pruning, our \PR\ algorithm enables us to perform $k$-object-sensitivity
analysis with a substantially much finer abstraction (larger $k$)
compared to a full $k$-object-sensitive analysis or even using the selected
refinement strategy of \cite{liang11minimal}.
In some cases, the non-pruning approaches hit a wall around $k=3$ but the \PR\
algorithm is able to go well beyond $k=10$.

### 2011-03-18 extra/sliver_pldi11/preliminaries.tex
fixed several typos

A {\em term} $t$ consists of a relation $t.r \in \sR$ and a list of arguments $t.\ba$,
where each argument $t.a_i$ is either a variable or a constant,
(that is, $t.a_i \in \sV \cup \sC$) for $i = 1, \dots, |t.\ba|$.
We will write a term in any of the following three equivalent ways:
\begin{align}
t \quad\equiv\quad t.r(t.\ba) \quad\equiv\quad t.r(t.a_1, \dots, t.a_{|t.\ba|}).
\end{align}
For example, $\ext(j,c,c')$ is a term.
We call a term whose arguments are all constants a {\em tuple}
(e.g., $\ext(\initNode, [\,], \vec{\initNode})$).
Note that the tuple includes the relation as well as the arguments.
We let $\xo$ denote a designated {\em query tuple} (e.g., $\common()$),
whose truth value we want to determine.

@@

% Explain
Condition (i) says that each tuple in a derivation
should either be given as an input tuple ($x_i \in X$)
or be the result of some rule $z \in \sZ$.
Condition (ii) says that $\xo$ was finally produced via some sequence of rule applications.
Condition (iii) says that the derivation of $x_n$ is minimal; each tuple was
used in some way; as we will see later, this condition is important because it allow us to
define the set of relevant tuples for pruning.

@@

% Example
In the context of our example (\reffig{graphExample}),
we could define an abstraction
that maps a sequence $c$ to the set of sequences that have the same first element as $c$
(e.g., $\alpha([10]) = \{ [1], [10], [11], \dots \}$).
Applied to a tuple, we get
\begin{align}
& \alpha(\ext(1, [0], [10])) = \\
& \quad\quad \ext(1, \{[0], [00], [01], \dots\}, \{[1], [10], [11], \dots\}). \nonumber
\end{align}
This abstraction is a simplified version of the abstractions we use in our
$k$-limited analyses (see \refsec{abstractions}).

% Extend to sets
We extend $\alpha$ to sets of constants (which we also call abstract values):
\begin{align}
\alpha(s) = \{ \alpha(c) : c \in s \}, \quad s \in \sP(\sC),
\end{align}
which returns a set of abstract values.
This allows us to naturally define compositions of abstractions
by flattening sets of abstract values:
\begin{align}
(\alpha_1 \circ \alpha_2)(c) \eqdef \cup_{s \in \alpha_1(\alpha_2(c))} s.
\end{align}
An important property of equivalence relations is
that if $\alpha_1 \preceq \alpha_2$, then $\alpha_1 \circ \alpha_2 = \alpha_1$
(applying a finer abstraction first has no impact).
Note that in general for arbitrary $\alpha_1,\alpha_2$,
the composition $\alpha_1 \circ \alpha_2$ is not an abstraction;
it is something that we must check when we consider compositions in
\refsec{abstractions}.

% Extend to tuples
We also extend $\alpha$ to concrete tuples $x$ and sets of concrete tuples $X$:
\begin{align}
\alpha(x) &= x.r(\alpha(x.a_1), \dots, \alpha(x.a_{|x.\ba|})), \\
\alpha(X) &= \{ \alpha(x) : x \in X \}.
\end{align}
Here, $\alpha(x)$ is an abstract tuple (one where the arguments are abstract values)
and $\alpha(X)$ is a set of abstract tuples.
Finally, we extend $\alpha$ to abstract tuples $b$ and sets of abstract tuples $B$:
\begin{align}
\alpha(b) &= \{ b.r(s_1, \dots, s_{|b.\ba|}) : \forall i, s_i \in \alpha(b.a_i) \}, \label{eqn:alphab} \\
\alpha(B) &= \cup_{b \in B} \alpha(b). \label{eqn:alphaB}
\end{align}
\refeqn{alphab} applies the abstraction function to each component and takes the cross product over the possible abstract values,
yielding a set of abstract tuples;
\refeqn{alphaB} aggregates these sets of abstract tuples.

@@

\FigStar{figures/graphDerivation}{0.3}{graphDerivation}{
Computation of $\bP(X)$ on the graph example from \reffig{graphExample}
under an abstraction which maps a path onto the set of paths with the same first element
(e.g., $\alpha(\vec{10}) = \{\vec{1}, \vec{10}, \vec{11}, \dots \} \eqdef \vec{1}*$).
Each abstract tuple is derived by a rule whose source terms corresponding to the incoming edges.
Relevant input tuples ($\bP(X)$, shown in green) are the ones which are reachable by following the edges backwards;
ones which are not are pruned ($X\backslash\bP(X)$, shown in red).
}

>>>>

A {\em term} $t$ consists of a relation $t.r \in \sR$ and a list of arguments $t.\ba$,
where each argument $t.a_i$ is either a variable or a constant,
(that is, $t.a_i \in \sV \cup \sC$) for $i = 1, \dots, |t.\ba|$.
We will write a term in any of the following three equivalent ways:
\begin{align}
t \quad\equiv\quad t.r(t.\ba) \quad\equiv\quad t.r(t.a_1, \dots, t.a_{|t.\ba|}).
\end{align}
For example, $\ext(j,c,c')$ is a term.
We call a term whose arguments are all constants a {\em tuple}
(e.g., $\ext(\initNode, [\,], \vec{\initNode})$).
Note that the tuple includes the relation as well as the arguments.
We let $\xo$ denote a designated {\em query tuple} (e.g., $\common(\Ga,\Gb,3)$),
whose truth value we want to determine.

@@

% Explain
Condition (i) says that each tuple in a derivation
should either be given as an input tuple ($x_i \in X$)
or be the result of some rule $z \in \sZ$.
Condition (ii) says that $\xo$ was finally produced via some sequence of rule applications.
Condition (iii) says that the derivation of $x_n$ is minimal, that is, each tuple was
used in some way; as we will see later, this condition is important because it allow us to
define the set of relevant tuples for pruning.

@@

% Example
In the context of our example (\reffig{graphExample}),
we could define an abstraction
that maps a sequence $c$ to the set of sequences that have the same first element as $c$
(e.g., $\alpha([10]) = \{ [1], [10], [11], \dots \}$).
This abstraction is a simplified version of the abstractions we use in our
$k$-limited analyses (see \refsec{abstractionExamples}).

% Extend to sets
We extend $\alpha$ to sets of constants (which we also call abstract values):
\begin{align}
\alpha(s) = \{ \alpha(c) : c \in s \}, \quad s \in \sP(\sC),
\end{align}
which returns a set of abstract values.
This allows us to naturally define compositions of abstractions
by flattening sets of abstract values:
\begin{align}
(\alpha_1 \circ \alpha_2)(c) \eqdef \cup_{s \in \alpha_1(\alpha_2(c))} s.
\end{align}
An important property of equivalence relations is
that if $\alpha_1 \preceq \alpha_2$, then $\alpha_1 \circ \alpha_2 = \alpha_1$
(applying a finer abstraction first has no impact).
Note that in general for arbitrary $\alpha_1,\alpha_2$,
the composition $\alpha_1 \circ \alpha_2$ is not an abstraction;
it is something that we must check when we consider compositions in
\refsec{algorithm}.

% Extend to tuples
We also extend $\alpha$ to concrete tuples $x$ and sets of concrete tuples $X$:
\begin{align}
\alpha(x) &= x.r(\alpha(x.a_1), \dots, \alpha(x.a_{|x.\ba|})), \\
\alpha(X) &= \{ \alpha(x) : x \in X \}.
\end{align}
Here, $\alpha(x)$ is an abstract tuple (one where the arguments are abstract values)
and $\alpha(X)$ is a set of abstract tuples.
For example:
\begin{align}
& \alpha(\ext(1, [0], [10])) = \\
& \quad\quad \ext(1, \{[0], [00], [01], \dots\}, \{[1], [10], [11], \dots\}). \nonumber
\end{align}
Finally, we extend $\alpha$ to abstract tuples $b$ and sets of abstract tuples $B$:
\begin{align}
\alpha(b) &= \{ b.r(s_1, \dots, s_{|b.\ba|}) : \forall i, s_i \in \alpha(b.a_i) \}, \label{eqn:alphab} \\
\alpha(B) &= \cup_{b \in B} \alpha(b). \label{eqn:alphaB}
\end{align}
\refeqn{alphab} applies the abstraction function to each component and takes the cross product over the possible abstract values,
yielding a set of abstract tuples;
\refeqn{alphaB} aggregates these sets of abstract tuples.

@@

\FigStar{figures/graphDerivation}{0.3}{graphDerivation}{
Computation of $\bP(X)$ on the graph example from \reffig{graphExample}
under an abstraction which maps a path onto the set of paths with the same first element
(e.g., $\alpha(\vec{10}) = \{\vec{1}, \vec{10}, \vec{11}, \dots \} \eqdef \vec{1}*$).
Each abstract tuple is derived by a rule whose source terms correspond to the incoming edges.
Relevant input tuples ($\bP(X)$, shown in green) are the ones which are reachable by following the edges backwards;
ones which are not are pruned ($X\backslash\bP(X)$, shown in red).
}
